---
title: "Chapter 3 Basic Syntax Report for Book *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics* (Week 2 Monday-)"
author: "Nurrospody"
date: "4/6/2020"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still in Progress, this currently has a version in SOURCE but not REPORTS (it will be in REPORTS once it is finished).  'dog' is placeholder vocabulary definition because otherwise my formatting breaks; many of these vocabulary terms are so simple that explaining them is a bit difficult, so I'll spend time doing that once I've fihnished the rest of the chapter.
  
### Notable Vocabulary 
**CRISP-DM** - dog  
**R Objects** - dog  
**Variables** - AKA. Variable is variable (python)  
**Data Types** - dog  
**Lists** - dog  
**Vectors** - A data structure for storing a SET of data values of the same type.  The most basic and common data structure.  
**Matrices** - dog  
**Data Frames** - dog  
**Conditional Statements** - dog  
**Loops** - dog  
**Functions** - dog  
**<-** - "Assign"  

Data Types | Example Values
-----------|---------------
Logical (Boolean)    | TRUE/FALSE (CAPS sensitive command)
Numeric           | 1(integer), 1.0(double), 0.33(double)
Character | "This is a bat" (quotes are mandatory)


### Writing in R Console - Basic Mathematics *Comands Used* : subraction, addition, division, multiplication, cosine, sine, tangent, standard deviation, mean, min, max, plot
I discovered that if I use semi-colons to mark the end of a line delimiter but DON'T break the line wtih ENTER, it compiles the information in a way that is more legible for an RMarkdown report.
``` {r math}
47-50; 89+6; 5 * 8; 54/870

cos(96); sin(96); tan(96)
```
Before solving for the standard deviation as-per the book's example, I noticed that a 'c' was part of the formula, but wasn't sure why.  I experimented a bit with not including it, and realized it only took the first value without c; c appears to be 'concatenate', to 'combine' something into a Vector; a list.  The help() page called it a 'Primitive' function.
```{r math contined}
sd(c(7,81,4,53,82,11,90, na.rm = FALSE)); mean(c(111, 65, 78, 93, 299)); min(c(9, 4, 7, 2)); max(c(9, 7, 2, 5))

plot(c(77, 45, 99, 88, 33))
``` 
 
 
### Making a basic script
Using information from the sections 'Using the Code Editor' and 'Adding Comments to the Code' I made the following simple RScript.  #Can be used to add a comment to a script, which has quickly become a favourite thing for me in this report!  The first part of the script assigns variables and then some math uses those assigned variables.  
I discovered that you can use = instead of <-, but apparently <- works on all levels of coding and = only works in the first level of code?  Not sure what code levels mean yet, but using <- for R seems generally god practice.
```{r simple script}
#Create Variable "A" with a value of "1"
A <- 1;
#Create Variable "B" with a value of "2"
B <- 2;

#Do mathematical functions with variables "A" & "B" (division, multiplication, addition, subtraction)
A/B; A * B; A + B; A - B;

#Square variables "A" and "B" on two seperate lines
A^2; B^2;
```


### Data Types *Commands Used* - (class), (is.datatype), (as.datatype)
In case the data type for a variable isn't known for some reason, (class) or (is.datatype) can be used.  Boolean can not be found as a function. "datatype" in is.datatype gets replaced by any of the types of data.
```{r cont}
#Determine the data type of a variable
print(class(A)); print(is.numeric(A)); print(is.character(A)); print(is.logical(A))
```
In specific cases, datatypes can be converted, such as the following case:
```{r data}
C <- 7;
D <- "7"; 
```
C + D results in a Line 74 Error in C + D : non-numberic argument to binary operator Calls and halts the Excecution.  D must be used as.numeric:
```{r as.type}
D <- as.numeric(D);
C + D;
```

### Vectors *Commands Used* c(), typeof(), class(), length(), : operator, print(), [] brackets for retrieving vectors, modify a vector with ->
```{r vector}
#Making a new variable, E, as a vector, then investigating it
E <- c(5,6,7,8,9); print(E); typeof(E); class(E); length(E);
#Making another new variable, F, as another larger vector for sokme additional excercises
F <- 4:20; print(F);
#What is the fifth value inside of this vector?  What about the fifth AND ninth?  All the values without the fifth?
F[5]; F[c(5,9)]; F[-5];
#Using E again, to test a shorter logical excercise: there are only 5 values in E, but I typed out 6 logical values, so the latst one is NA.
E[c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE)];
#Back to Variable F.  Let's test using the less than sign to retrieve an element from the vector.  There is nothing <4, but there is something <5.
F[F<4]; F[F < 5];
#Let's look at E one last time.  What happens if we modify a vector?  If we modify the third value, it gets REPLACED, not ADDED IN.
print(E);
E[3] <- 9;
print(E);
```


