---
title: "Chapter 3 Basic Syntax Report for Book *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics* (Week 2 Monday-)"
author: "Nurrospody"
date: "4/6/2020"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still in Progress, this currently has a version in SOURCE but not REPORTS (it will be in REPORTS once it is finished).  'dog' is placeholder vocabulary definition because otherwise my formatting breaks; many of these vocabulary terms are so simple that explaining them is a bit difficult, so I'll spend time doing that once I've fihnished the rest of the chapter.
  
### Notable Vocabulary 
**CRISP-DM** - dog  
**R Objects** - dog  
**Variables** - AKA. Variable is variable (python), something that is defined and then can be used in the console.  
**Data Types** - What a data's classification is--numeric, character, TRUE/FALSE.  
**Lists** - Like a vector, a list can store a set of values, but it can store values of multiple data types in a single list.    
**Vectors** - A data structure for storing a SET of data values of the same type.  The most basic and common data structure.  
**Matrix** - Like a vector, but in a table--a vector with 'two dimensions' for the data.  All data must of the same type.    
**Data Frames** - Usually stores data from EXCEL or .csv. these are like matrices that contain multiple data types.  They may look like charts with both 'names' and numbers to go with those 'names'.  
**Conditional Statements** - dog  
**Loops** - dog  
**Functions** - dog  
**<-** - "Assign"  

Data Types | Example Values
-----------|---------------
Logical (Boolean)    | TRUE/FALSE (CAPS sensitive command)
Numeric           | 1(integer), 1.0(double), 0.33(double)
Character | "This is a bat" (quotes are mandatory)


### Writing in R Console - Basic Mathematics *Comands Used* : subraction, addition, division, multiplication, cosine, sine, tangent, standard deviation, mean, min, max, plot
I discovered that if I use semi-colons to mark the end of a line delimiter but DON'T break the line wtih ENTER, it compiles the information in a way that is more legible for an RMarkdown report.
``` {r math}
47-50; 89+6; 5 * 8; 54/870

cos(96); sin(96); tan(96)
```
Before solving for the standard deviation as-per the book's example, I noticed that a 'c' was part of the formula, but wasn't sure why.  I experimented a bit with not including it, and realized it only took the first value without c; c appears to be 'concatenate', to 'combine' something into a Vector; a list.  The help() page called it a 'Primitive' function.
```{r math contined}
sd(c(7,81,4,53,82,11,90, na.rm = FALSE)); mean(c(111, 65, 78, 93, 299)); min(c(9, 4, 7, 2)); max(c(9, 7, 2, 5))

plot(c(77, 45, 99, 88, 33))
``` 
 
 
### Making a basic script
Using information from the sections 'Using the Code Editor' and 'Adding Comments to the Code' I made the following simple RScript.  #Can be used to add a comment to a script, which has quickly become a favourite thing for me in this report!  The first part of the script assigns variables and then some math uses those assigned variables.  
I discovered that you can use = instead of <-, but apparently <- works on all levels of coding and = only works in the first level of code?  Not sure what code levels mean yet, but using <- for R seems generally god practice.
```{r simple script}
#Create Variable "A" with a value of "1"
A <- 1;
#Create Variable "B" with a value of "2"
B <- 2;

#Do mathematical functions with variables "A" & "B" (division, multiplication, addition, subtraction)
A/B; A * B; A + B; A - B;

#Square variables "A" and "B" on two seperate lines
A^2; B^2;
```


### Data Types *Commands Used* - (class), (is.datatype), (as.datatype)
In case the data type for a variable isn't known for some reason, (class) or (is.datatype) can be used.  Boolean can not be found as a function. "datatype" in is.datatype gets replaced by any of the types of data.
```{r cont}
#Determine the data type of a variable
print(class(A)); print(is.numeric(A)); print(is.character(A)); print(is.logical(A))
```
In specific cases, datatypes can be converted, such as the following case:
```{r data}
C <- 7;
D <- "7"; 
```
C + D results in a Line 74 Error in C + D : non-numberic argument to binary operator Calls and halts the Excecution.  D must be used as.numeric:
```{r as.type}
D <- as.numeric(D);
C + D;
```

### Vectors *Commands Used* c(), typeof(), class(), length(), : operator, print(), [] brackets for retrieving vectors, modify a vector with ->
```{r vector}
#Making a new variable, E, as a vector, then investigating it
E <- c(5,6,7,8,9); print(E); typeof(E); class(E); length(E);
#Making another new variable, F, as another larger vector for sokme additional excercises
F <- 4:20; print(F);
#What is the fifth value inside of this vector?  What about the fifth AND ninth?  All the values without the fifth?
F[5]; F[c(5,9)]; F[-5];
#Using E again, to test a shorter logical excercise: there are only 5 values in E, but I typed out 6 logical values, so the latst one is NA.
E[c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE)];
#Back to Variable F.  Let's test using the less than sign to retrieve an element from the vector.  There is nothing <4, but there is something <5.
F[F<4]; F[F < 5];
#Let's look at E one last time.  What happens if we modify a vector?  If we modify the third value, it gets REPLACED, not ADDED IN.
print(E);
E[3] <- 9;
print(E);
```

### Lists *Non-duplicate Commands Used* list(), str(), [[sublist]], NULL  
This section used a lot of duplicate commands from the Vector section, so only the new ones are shown.
```{r lists}
G <- list("fox", "phoenix", "water", 9); print(G); str(G);
#There seems to be a difference between a list and a sublist; I'll have to figure the practical use of this difference with supplementary material.  
G[1]; G[[1]];
#NULL (caps sensitive?) is a new command for assigning NULL to part of a list, I presume it would work for a vector as well..
G[[4]] <- NULL; print(G)
```

### Matrix *Non-Duplicate Comands Used* matrix(), nrow, ncol, dimnames, attributes(), colnames(), rownames(), cbind(), rbind(), t()
```{r matrix}
#To experiment with making matrices, I made up a a pretend survey where 6 people were asked if they liked or disliked 3 different animals.  Making this was clunky, however . . . there must be better ways to make tables and better uses of matrices.  Row is the X axis and Column is the Y axis.  
#Entering the data goes down column one and then loops around to column two, instead of across row one and then looping to row two.
H <- matrix(c(1,4,5,5,2,1), nrow=3, ncol=2, dimnames=list(c("Spider", "Fox", "Dog"), c("Likes the Animal", "Dislikes the Animal"))); print(H)
#These are some commands to investigate the matrix.
attributes(H); colnames(H); rownames(H);
#These comands use row and column binding to make matrices; column binding makes the values travel down and then loop to the next column; row binding makes the values travel across the rows and then loop down to the next column.
I <- cbind(c(1,2,3), c(4,5,6)); print(I); J <- rbind(c(1,2,3), c(4,5,6)); print(J);
#Selecting different rows & columns in a matrix works similar to a vector, but if the comma is before the number it selects the relevant column.  If the comma is AFTER the number it selects the relevant row.
print(J); J[,1]; J[2,]; J[-1,]; J[2,2]
#You can also add an additional row or column to a matrix by using rbind or cbind respectfully; the technique taught in the book technically makes a whole new variable.
K <- cbind(c("old", "old"), c("old", "old")); print(K); L <- rbind(K, c("new", "new")); print(L)
#I wanted to practice modifying a matrix: I think the order of [column, row] confuses me since it's going Y and then X for reading,  
#instead of X and then Y, so it's the inverse of reading a graph (kind of, this isn't meant to analogy with a graph, but figuring out why I was confused is helpful.)
J[2,3] <- 66; print(J);
#With t() matrixes can be transposed--i.e. which axis each label is on is switched. 
print(L); L <- t(L);print(L)
```
### Data Frames *Commands Used* data.frame()  
I tried making an initial data.frame that had number of legs (0, 4, 8), animal name (snake, fox, spider), and Does Riley like it? (good, good, AAAAAAAA) but it proved to be quite difficult; I kept getting errors about unexpected commas, or unexpected =.  I had to look around for a while.  Note to self: I might take what I have right now and bring it to the QUASR and instead figure out how to import a .csv and work with that instead to practice.
```{r dataframe}
#Making this silly dataframe was more difficult than it should have been.  Extra =s, extra commas. . .
#M <- data.frame(number_of_legs(c(0, 4, 8)), animal(c("Snake", "Fox", "Spider")), Does_Riley_Like_it?(c(("good", "good", "AAAAAAAA")); print(M);
```
