---
title: "Chapter 4 part4 *Descriptive Statistics* (Topics Go Here)"
author: "Nurrospody"
date: "5/5/2020, *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics*"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Central Tendency, Spread, Variation Commands that DO require manual help CONTINUED

In the last report we ended by talking about getting the mode of a dataset or variable, and how that requires making a special function or frequency tahble.  
Next we'll talk about getting **POPULATION** (POP) **Standard Deviation and Variance**, which also require 'special' treatment.  

Because of Bessel's correction--which is meant to partially correct bias of standard deviation estimation--we need to append ```* (N - 1)/N:``` to our variance var() function.  N is length(variable), or the amount of points in our variable.  N-1 is used for similar reasons as n-1 in other statistical formulae.   
In order to get the POP Standard Deviation, we do the above the get the POP variance, and then sqrt(variance).  We do this, rather than ```sd(X) * (N-1) / N;```, because Variance is a better unbiased estimator than Standard Deviation.  So in order to get the POP Standard Deviation in R, you must get the POP Variance first.

Let's calculate the POP Variance and Standard deviation using our 'random' variable.
```{r random setup, include=FALSE}
random <- rnorm(70, 10, 2);
```
```{r ran}
N <- length(random);
variance <- var(random) * (N - 1) / N;
N; variance; sqrt(variance);
```
The first number in the solution box is N (70), the second number is the POP Variance (probably about 4), and the third number is the POP Standard Deviation (probably about 2).  Because Standard Deviation is the square of variance, it of course has a less wide range than variance does.  I guess that inherently makes it more biased because it is smaller and more concentrated.  






To continue reading the CH4 reports, select a new section:  
[Part 5 of the Chapter 4 Reports *Will result in error page currently*](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/Chapter%20Reports/CH4-part5.md)  
[Link to README to select any Chapter Report](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/README.md)