---
title: "CH6 part 2 *Inferential Statistics & Regressions*"
author: "Nurrospody"
date: "7/24/2020, *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics*"
output:
  github_document:
    toc: TRUE
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Modes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

Madoka <- read.csv(file="../data_sources/Madoka Magika Volume 1-3.csv")

Smallmad <- Madoka[1:5,]
require(dplyr)

Crop <- read.csv(file="../data_sources/crop.data.csv")
require(AICcmodavg)
require(ggplot2)

require(xlsx)
Garden <- read.xlsx(file="../data_sources/Garden Practicum.xlsx", 1)
Garden2 <- read.xlsx(file="../data_sources/Garden Practicum.xlsx", 2)
```

*Let's revisit some sections that only had basic notes written down in Chapter6 part1.*  

### T-Test 187-194    

#### **1 sample t-test**
Question: is the mean of a population equal to ___?  
```t.test()```  function  
When gathering data, we don't actually know what the mean of a whole population is.  But we can use a 1-way T-test to figure out if a population mean really could be something, not based on random chance alone.  
The following isn't really a good application for this test, but at least we can do it.  
```{r}
t.test(Crop$yield, mu=100)
```
Here, we're testing for if the mean of "Crop" 'yield' variable is equal to 100.  
The probability that the real mean is actually 100, with the data we've given, is obscenely low.  This could mean a couple of things.  1. there's no way the mean is 100 (which is the case here) OR the experiment that found a mean that wasn't equal to 100 used a non-random sampling method.  
A 1-way T-test is how the ideda that a human's mean body temperature was 98.6 was disproved.  I would have used that dataset for this example, but it's protected from use except for classes that are using the Triola textbook.  (I own that book, but this ILC technically doesn't use it.)  

#### **2 sample t-test**
Question: are the means of these two samples the same?  
I needed two datasets that were similar for this one.  Even though the sample size is not quite big enough yet, I decided to load in the [garden dataset that I've been making](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/data_sources/Garden%20Practicum.xlsx).  I'll do a more comprehensive look into it later, so it's fine.  I loaded Sheet 1.
```{r}
t.test(Garden$Pansy2, Garden$Pansy5, var.equal=TRUE, paired=FALSE)
```
I chose to look at pansy 2 and pansy 5 because they have been very succesful at different times in the past 26 days.  
The rest results tell us that the probability that these means are the same is 0.4%, approximately.  This is much lower than alpha, 0.05 (5%), so we reject the null hypothesis (that they are the same).  We get the added benefit of seeing what the means are at the bottom, so we can actually tell which mean is higher than the other without having to do extra postHoc tests like with ANOVA.  If I wanted to test all 5 pansies at once, I could use a 1-way ANOVA, but that's a little bit unnesseary since I'm interested in sample mean, not variance.  Still, it'd be simpler than running 20 different 2-ways.  

### Chi-Square Test 194-198  
relationships between categorical variables/data (rather than continuous data)  
uses hypothesis test null/alternate arguments.  

#### **One categorical variable (Goodness of Fit)**  
Question: Is this one sample we have, consistent with the distribution of the other ones?  
```chisq.test(data)```  function 

```{r}
```



#### **Two categorical variable (Contingency Test)** 
Question: is there a relationship between these two variables?  
```chisq.test(data.table)``` function  
I decided to see if the distribution of live pansies, was consistant with the distribution of dead pansies.  The amount of dead pansies s less than live ones on any given days, but will it see a similar pattern?  I need to review what chi-square tests are in my basic stats textbook, so seeing the result should help.  
```{r}
Pansylife <- data.frame(Garden$Pansy3, Garden2$Pansy3);
Pansytable <- table(Pansylife);
Pansytable
```
. . . I forgot that it had to be categorical variables and not numerical, so this result doesn't make a lot of sense.  At least that definitely shows what a chi-square test is supposed to be for!!
I was thinking of a Correlation test ```cor()``` which I still have to revisit.  

Making mistakes about what tests do what has actually been super helpful so far for figuring out what does what, though.
