---
title: "CH6 part4 Linear Regressions"
author: "Nurrospody"
date: "8/17/2020, *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics*"
output:
  github_document:
    toc: TRUE
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Modes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

Madoka <- read.csv(file="../data_sources/Madoka Magika Volume 1-3.csv")

Smallmad <- Madoka[1:5,]
require(dplyr)

Crop <- read.csv(file="../data_sources/crop.data.csv")
require(AICcmodavg)
require(ggplot2)

require(xlsx)
Garden <- read.xlsx(file="../data_sources/Garden Practicum.xlsx", 1)
Garden2 <- read.xlsx(file="../data_sources/Garden Practicum.xlsx", 2)
require(moments)
FB <- read.xlsx(file="../data_sources/FriendshipBonus.xlsx", 1)
SleepyFB <- read.xlsx(file="../data_sources/PCM_Pokemon_Visitors_Data_Data.xlsx", 2)
library(random)
```

### Linear Regression (lm)  
With a linear regression, we can test for a linear relationship between one independent variable and one dependent variable.  
I'm going to Arc's friendship bonus data, and test for if ```FB$Stage``` (independent variable) has any linear relationship with ```FB$StagesSince``` (dependent variable).  The research question is: If the Stage number is higher, do friendship bonuses occur more often?  
The null hypothesis is that there is no linear relationship; the alternate hypothesis is that the number of a stage has a statistically significant effect on the stages since a friendship bonus was received.  

#### Testing for normality  
In a previous excercise, I had already used ```shapiro.test(FB$StagesSince[2:31])``` to test for normality (I could not find normality, presumably because of n30); I had also already used ```skewness(FB$StagesSince[2:31])``` to test for normality of skew.  
However, this time I will instead use a histogram and linear plot to test for linearity.  
```{r}
hist(FB$StagesSince); plot(Stage ~ StagesSince, data = FB)
```

The histogram appears to be relatively normal; however, we can see with the basic plot that no linear relationship exists.  **We can immediately fail to reject the null hypothesis**.  

For the sake of this excercise, though, I will complete the linear analysis so that I can use the commands.  

```{r}
#lm(Dependent variable ~ independent variable)
StageNum.StagesSince.lm <- lm(StagesSince ~ Stage, data = FB)
summary(StageNum.StagesSince.lm)
```
In the summary, we see our model equation, standard error, t-value, and p-value.  
(Intercept) is the y-intercept of the regression equation, but because there is no linear relationship it doesn't matter here.  Othewise, we could use this data to predict where points that we don't have may fall within our range of variables.  (Predicting outside sometimes works, but can also be completely wrong, such as with income/happiness data.)  

Our p-value for Stage is 0.306--much greater than alpha 0.05.  We fail to reject our null hypothesis that there is no linear relationship.


#### Spearman’s rank correlation test  
This is the non-parametric test that can be run for data that does not fulfil the normality requirements (or similar variance requirements aka homoscedasticity) for the parametric linear regression test.  
We would need to ```install.packages("ggpubr")```, load the library ```library("ggpubr")```, and then use the function ```cor.test()``` rather than lm() which we used for the parametric test.  I have inferred from the gg- in front of the package name that this is related to the ggplot package!

### Multiple Linear Regression  
With multiple linear regression, we can test for linear relationships between multiple independent variables and one dependent variable.



To continue reading the CH6 reports, select a new section:  
Next: [Part 5 of the Chapter 6 Reports](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/Chapter%20Reports/CH6-part5.md)  
Previous: [Part 3 of the Chapter 6 Reports](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/Chapter%20Reports/Ch6-part3.md)   
  
[Link to README to select any Chapter Report](https://github.com/Nurrospody/SOURCE-Statistics-ILC/blob/master/README.md)