---
title: "Chapter 3 Basic Syntax Report for Book *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics* (Started Week 2 Monday)"
author: "Nurrospody"
date: "4/6/2020"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still in Progress.  
'dog' is placeholder vocabulary definition because otherwise my formatting breaks; many of these vocabulary terms are so simple that explaining them is a bit difficult, so I'll spend time doing that once I've fihnished the rest of the chapter.
  
### Notable Vocabulary 
**CRISP-DM** - dog  
**R Objects** - dog  
**Variables** - AKA. Variable is variable (python), something that is defined and then can be used in the console.  
**Data Types** - What a data's classification is--numeric, character, TRUE/FALSE.  
**Vectors** - A data structure for storing a SET of data values of the
SAME type. The most basic and common data structure. If stored with
extra attributes, like dim, it is technically an ‘array’ and not a
vector.  
**Lists** - Like a vector, a list can store a SET of values, but
it can store values of DIFFERENT data types in a single list.  
**Matrix** - Stores data in a ‘two dimensional’ table rather than a
one-dimensional list. All data must of the SAME type. Uses dimnames to
label each axis of the table.  
**Data Frames** - Stores data in a ‘two dimensional’ table; can be of
DIFFERENT types. Usually stores data imported from EXCEL or .csv. Uses
dimnames to label axes of the table.  
**Conditional Statements** - dog  
**Loops** - dog  
**Functions** - dog  
**<-** - "Assign"  

Data Types | Example Values
-----------|---------------
Logical (Boolean)    | TRUE/FALSE (CAPS sensitive command)
Numeric           | 1(integer), 1.0(double), 0.33(double)
Character | "This is a bat" (quotes are mandatory)


### Writing in R Console - Basic Mathematics *Comands Used* : subraction, addition, division, multiplication, cosine, sine, tangent, standard deviation, mean, min, max, plot
I discovered that if I use semi-colons to mark the end of a line delimiter but DON'T break the line wtih ENTER, it compiles the information in a way that is more legible for an RMarkdown report.
``` {r math}
47-50; 89+6; 5 * 8; 54/870

cos(96); sin(96); tan(96)
```
Before solving for the standard deviation as-per the book's example, I noticed that a 'c' was part of the formula, but wasn't sure why.  I experimented a bit with not including it, and realized it only took the first value without c; c appears to be 'concatenate', to 'combine' something into a Vector; a list.  The help() page called it a 'Primitive' function.
```{r math contined}
sd(c(7,81,4,53,82,11,90, na.rm = FALSE)); mean(c(111, 65, 78, 93, 299)); min(c(9, 4, 7, 2)); max(c(9, 7, 2, 5))

plot(c(77, 45, 99, 88, 33))
``` 
 
 
### Making a basic script
Using information from the sections 'Using the Code Editor' and 'Adding Comments to the Code' I made the following simple RScript.  #Can be used to add a comment to a script, which has quickly become a favourite thing for me in this report!  The first part of the script assigns variables and then some math uses those assigned variables.  
I discovered that you can use = instead of <-, but apparently <- works on all levels of coding and = only works in the first level of code?  Not sure what code levels mean yet, but using <- for R seems generally god practice.
```{r simple script}
#Create Variable "A" with a value of "1"
A <- 1;
#Create Variable "B" with a value of "2"
B <- 2;

#Do mathematical functions with variables "A" & "B" (division, multiplication, addition, subtraction)
A/B; A * B; A + B; A - B;

#Square variables "A" and "B" on two seperate lines
A^2; B^2;
```


### Data Types *Commands Used* - (class), (is.datatype), (as.datatype)
In case the data type for a variable isn't known for some reason, (class) or (is.datatype) can be used.  Boolean can not be found as a function. "datatype" in is.datatype gets replaced by any of the types of data.
```{r cont}
#Determine the data type of a variable
print(class(A)); print(is.numeric(A)); print(is.character(A)); print(is.logical(A))
```
In specific cases, datatypes can be converted, such as the following case:
```{r data}
C <- 7;
D <- "7"; 
```
C + D results in a Line 74 Error in C + D : non-numberic argument to binary operator Calls and halts the Excecution.  D must be used as.numeric:
```{r as.type}
D <- as.numeric(D);
C + D;
```

### Vectors *Commands Used* c(), typeof(), class(), length(), : operator, print(), [] brackets for retrieving vectors, modify a vector with ->
```{r vector}
#Making a new variable, E, as a vector, then investigating it
E <- c(5,6,7,8,9); print(E); typeof(E); class(E); length(E);
#Making another new variable, F, as another larger vector for sokme additional excercises
F <- 4:20; print(F);
#What is the fifth value inside of this vector?  What about the fifth AND ninth?  All the values without the fifth?
F[5]; F[c(5,9)]; F[-5];
#Using E again, to test a shorter logical excercise: there are only 5 values in E, but I typed out 6 logical values, so the latst one is NA.
E[c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE)];
#Back to Variable F.  Let's test using the less than sign to retrieve an element from the vector.  There is nothing <4, but there is something <5.
F[F<4]; F[F < 5];
#Let's look at E one last time.  What happens if we modify a vector?  If we modify the third value, it gets REPLACED, not ADDED IN.
print(E);
E[3] <- 9;
print(E);
```

### Lists *Non-duplicate Commands Used* list(), str(), [[sublist]], NULL  
This section used a lot of duplicate commands from the Vector section, so only the new ones are shown.
```{r lists}
G <- list("fox", "phoenix", "water", 9); print(G); str(G);
#There seems to be a difference between a list and a sublist; I'll have to figure the practical use of this difference with supplementary material.  
G[1]; G[[1]];
#NULL (caps sensitive?) is a new command for assigning NULL to part of a list, I presume it would work for a vector as well..
G[[4]] <- NULL; print(G)
```

### Matrix *Non-Duplicate Comands Used* matrix(), nrow, ncol, dimnames, attributes(), colnames(), rownames(), cbind(), rbind(), t()
```{r matrix}
#To experiment with making matrices, I made up a a pretend survey where 6 people were asked if they liked or disliked 3 different animals.  Making this was clunky, however . . . there must be better ways to make tables and better uses of matrices.  Row is the X axis and Column is the Y axis.  
#Entering the data goes down column one and then loops around to column two, instead of across row one and then looping to row two.
H <- matrix(c(1,4,5,5,2,1), nrow=3, ncol=2, dimnames=list(c("Spider", "Fox", "Dog"), c("Likes the Animal", "Dislikes the Animal"))); print(H)
#These are some commands to investigate the matrix.
attributes(H); colnames(H); rownames(H);
#These comands use row and column binding to make matrices; column binding makes the values travel down and then loop to the next column; row binding makes the values travel across the rows and then loop down to the next column.
I <- cbind(c(1,2,3), c(4,5,6)); print(I); J <- rbind(c(1,2,3), c(4,5,6)); print(J);
#Selecting different rows & columns in a matrix works similar to a vector, but if the comma is before the number it selects the relevant column.  If the comma is AFTER the number it selects the relevant row.
print(J); J[,1]; J[2,]; J[-1,]; J[2,2]
#You can also add an additional row or column to a matrix by using rbind or cbind respectfully; the technique taught in the book technically makes a whole new variable.
K <- cbind(c("old", "old"), c("old", "old")); print(K); L <- rbind(K, c("new", "new")); print(L)
#I wanted to practice modifying a matrix: I think the order of [column, row] confuses me since it's going Y and then X for reading,  
#instead of X and then Y, so it's the inverse of reading a graph (kind of, this isn't meant to analogy with a graph, but figuring out why I was confused is helpful.)
J[2,3] <- 66; print(J);
#With t() matrixes can be transposed--i.e. which axis each label is on is switched. 
print(L); L <- t(L);print(L)
```
### Data Frames *Non-duplicate Commands Used* data.frame(), read.csv(). *Investigated Escape Character; cmd; replacement weirdness*

I initially tried making a data.frame with the book-given command but it
proved to be quite difficult; I kept getting errors about unexpected
commas, or unexpected =. I will figure it out, but first I’ll read a
.csv into R to work with instead.

``` {r data.frame}
#M <- data.frame(number_of_legs(c(0, 4, 8)), animal(c("Snake", "Fox", "Spider")), Does_Riley_Like_it?(c("good", "good", "AAAAAAAA")); print(M);
#Also tried
#M <- data.frame(number_of_legs(c=0, 4, 8), animal(c="Snake", "Fox", "Spider"), Does_Riley_Like_it?(c="good", "good", "AAAAAAAA")); print(M);
```
I knew I had a few .csv files on my computer from previous statistics
projects and wanted to also learn how to find those, so I searched for
them with cmd.

``` {r cmd}
#where looks in the directory; /F formats the results to have quotes around them, /R Recursive looks into ALL directories, * is a wildcard that can include anything before the .csv result.
#C:\Users\Persimmon>where /F /R %userprofile% *.csv
```

I found 8 .csv files scattered across my documents and OneDrive, and
chose one that I had made to take a single full-soundtrack video and
turn that into seperated mp3s. I had to copy the file from my OneDrive
into my Documents folder to open it properly in the Console, and I had
to make a data\_sources folder in my repository to get it to open in the
Rmd file. I discovered, also, that / and \\ work the same(ish) in
Windows devices, but \\ if typed alone is an Escape Character so using
DOUBLE \\ if using foward slash rather than backslash for addresses is
nessesary.

``` {r Madoka}
Madoka <- read.csv(file="../data_sources/Madoka Magika Volume 1-3.csv"); print(Madoka);
```

For some reason when I looked at the structure or nrow ncol of Madoka I
got 32 rows and 10 columns, but when I looked at Madoka\[0\] I got a
result that seemed to claim there were no columns in the dataset.

``` {r Madoka2}
str(Madoka);ncol(Madoka);nrow(Madoka); Madoka[0]; Madoka[,0];
```

I didn’t want to continue printing the whole Madoka data.frame with 10
columns and 32 rows when experimenting with different commands, so I
finagled with commands until I got the first 5 rows of the data.frame,
and then made that a new variable. I got a lot of hands-on experience
with trying to select a specific row/column of a variable here, doing it
with an actual .csv with many rows and columns took much more
understanding and exploration than just the small made-up data sets.

``` {r small}
Smallmad <- Madoka[1:5,]; print(Smallmad)
```

I then wanted to change one of the columns in Smallmad, but something
very perculiar happaned. Any input to [5,3] other than the original
input, “00:10:39” would result in an invalid factor level, and generate
an \<NA\>. If I changed it back to the original input, it worked
properly again.  
Not-pictured below: I thought that maybe the data format being ‘factor’
meant that it “factored” somehow with the other values. I don’t expect
that the program is smart enough to recognize that “xx:xx:xx” is a time
format without being specifically told that it is a time format, but I
still tried changing it so it didn’t technically contradict any
timestamps I could see. However, even changing [5,3] to “00:10:40”
resulted in \<NA\>.

``` {r small2}
Smallmad[5,3] <- "The angriest Madoka"; print(Smallmad); Smallmad[5,3] <- "00:10:39"; print(Smallmad);
```

Here I read ahead in the book a bit and realized how long the chapters are--having each chapter continue to be its report is a bad idea, so I'm ending this part of the report here.