---
title: 'Logical Statements & Loops : Chapter 3 *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics*'
author: "Nurrospody"
date: "4/21/2020"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

On Week 4, Tuesday, I experimented with loops and logical statements.  My book presented these ideas seperately, but they really make the most sense together.  
First, I made variable "pumpkin";
```{r define pumpkin}
pumpkin <- c(1:13); print(pumpkin)
```
Next, I needed to decide what to do with pumpkin.  In this case for each variable in pumpkin, I want to *loop* *for* if the elements and/or variables in pumpkin are divisible by 2. "p" stands for each variable in the vector "pumpkin".
```{r loop pumpkin}
for (p in pumpkin) {
    #for each element/value in pumpkin, print if that value is divisible by 2 
  }
```
This loop isn't useful on its own without a logical statement to pair with.  Our *if else* statment is that if the leftoever after dividing p by 2 is not 0 (modulus arithmetic), pumpkin does not divide into 2.  
```{r pumpkin logic}
for (p in pumpkin) {
    #	for each element/value in pumpkin, print if that value is divisible by 2
    if(p %% 2 != 0) { 
        message(p, " NO, pumpkin not divisible by 2");
        #cat(sprintf("%d is not divisible by 2\n", p))
    } else {
        message(p, " YES, pumpkin divisible by 2");
        #cat(sprintf("%d is\n", p));
    }
}
```
The comments after printing each message (the cat, sssss whatever s meant, print, format) is from when I was trying to figure out with my partner how to print each number that did or did not divide evenly with 2.

I kept one of my failed attempts, to explain what didn't work when making these (the whole thing will have to be a comment, otherwise Rmd won't knit correctly):
```{r fail}
#for (pumpkin in 1:13) {
     #for each element/value in pumpkin, print if that value is divisible by 2
 #}
 #if(pumpkin %% 2 != 0) { 
 #+ print("NO, pumpkin not divisible by 2");
     #+ } else {
     #+ print("YES, pumpkin divisible by 2");
```
In this attempt, the {} for the for loop didn't correctly close around the logical statement, so the loop didn't actually do anything; and I had been trying to divide pumpkin (a vector) instead of the variables inside of pumpkin (so I needed p, not pumpkin).  I also had not told it that I wanted each value that did/did not divide into 2 to be displayed, so when I did fix it, it simply showed [1] by each result which wasn't helpful.
