---
title: 'Matrices and Data.frame (Madoka.csv) : Chapter 3 *Learn R for Applied Statistics : With Data Visualizations, Regressions, and Statistics*'
author: "Nurrospody"
date: "Week 3 Wednesday-Friday"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Matrix *Non-Duplicate Comands Used* matrix(), nrow, ncol, dimnames, attributes(), colnames(), rownames(), cbind(), rbind(), t()

To experiment with matrices, I made up a a pretend survey where 6 people were asked if they liked or disliked 3 different animals.  Making this felt a bit clunky . . . either I need more practice, or there are better uses for matrices than tables like this.  
When entering the data, it fills in the table like this: the data goes all the way down column one and then loops around to column two.  I had to make sure when entering my data, that I wasn't thinking: how many people liked the spider; how many people disliked the spider? I had to think: how many people liked the spider, the fox, the dog?  Only once I'd filled in my first column completely, could I then think about the second column.
```{r matrix}

H <- matrix(c(1,4,5,5,2,1), nrow=3, ncol=2, dimnames=list(c("Spider", "Fox", "Dog"), c("Likes the Animal", "Dislikes the Animal"))); print(H)
#These are some commands to investigate the matrix.
attributes(H); colnames(H); rownames(H);
```
Selecting different rows & columns in a matrix works similar to a vector, but if the comma is before the number [,n] it selects the relevant column.  If the comma is AFTER the number it selects the relevant row. [n,].
``` {r selecting and editing}
print(H); H[,1]; H[2,]; H[-1,]; H[2,2]
```
This makes sense if you think about selecting position in a vector (or data.frame) to be like reading a graph (except it's reversed because the data goes down, not up).  Remember that a column's NAME is on the X axis, even though it then travels down the Y axis, and visa versa.  Where the name is, is what's important.
Because I had some trouble with this, I also practiced replacing a value inside of a matrix until I understood *where* in the matrix was being modified.
```{r editing}
#oh no!  Suddenly a bunch of dog-disliking people stormed in to our animal survey!  Now our results are not very useful, because this is obviously a biased sample.
H[3,2] <- 66; print(H);
```
We can add an additional row or column to a matrix by using rbind() or cbind() respectfully.  Technically, the command shown here makes an additional variable rather than modifying an already existing one.
```{r bind}
K <- cbind(c("old", "old"), c("old", "old")); print(K); L <- rbind(K, c("new", "new")); print(L)
```
With t() matrixes can be transposed--i.e. which axis each label is on is switched. 
```{r transpose}
print(L); L <- t(L);print(L)
```


#### Data Frames *Non-duplicate Commands Used* data.frame(), read.csv(). *Investigated Escape Character; cmd; replacement weirdness*

Initially I tried making a data.frame with the book-given command.  Eventually after many trials, giving up, and coming back to it a week later when I could actually copy paste the syntax from my book, I got it to work:
```{r Riley animal dataframe}
RilA <- data.frame(number_of_legs=c(0, 4, 8), animal=c("Snake", "Fox", "Spider"), Rileys_Opinion=c("good", "good", "AAAAAAAA")); print(RilA)
```
Originally I kept getting errors about unexpected commas, wrong ()s, or unexpected =, most of it was format issues.  Then I learned that I can't put a "?" or a "'" in a string name or the code will break, and originally the third column was "Does_Riley_Like_It?" or "Riley's Opinion" so that explains it, too.

I didn't have this data.frame when I actually did the excercises, so instead I took the more practical approach of loading a *.csv* file into R as a data.frame.

I had a few *.csv* files on my computer from previous statistics projects, and I learned how to find them using Command Line.  I've shown my search and what each part of the search means below as a comment:
``` {r cmd}
#where looks in the directory; /F formats the results to have quotes around them, /R Recursive looks into ALL directories, * is a wildcard that can include anything before the .csv result.
#C:\Users\Persimmon>where /F /R %userprofile% *.csv
```
I found 8 .csv files and chose one meant to take an hour-length soundtrack video and turn that into seperated mp3s. This not only taught me how to find and open a *.csv* but also about data pathing;  
I had to copy the file from my OneDrive into my Documents folder to open it properly in the RConsole; and in order to knit it from my R Code Editor, I had to place it inside of a data_sources subdirectory in my Git directory (where all my md and Rmd files are located).   
I also discovered that / and \\ work the same(ish) in Windows devices for pathing, but \\ if typed alone is an Escape Character.  Using DOUBLE \\ if using backslash rather than forwardslash is neccesary, for the sake of typing addresses or any markdown documents.

``` {r Madoka}
Madoka <- read.csv(file="../data_sources/Madoka Magika Volume 1-3.csv"); print(Madoka);
```

When I looked at the structure or nrow ncol of Madoka I got 32 rows and 10 columns, but when I looked at Madoka\[0\] I got a result that seemed to claim there were no columns in the dataset.  I thought this was interesting.

``` {r Madoka2}
str(Madoka);ncol(Madoka);nrow(Madoka); Madoka[0]; Madoka[,0];
```

I didn’t want to continue printing the whole Madoka data.frame with 10 columns and 32 rows when experimenting with different commands, so I took the first 5 rows of the data.frame and made that its own variable.
I got a lot of hands-on experience with trying to select a specific row/column of a variable here.  Doing it with an actual *.csv* with many rows and columns took more understanding and exploration than just the small made-up data sets.

``` {r small}
Smallmad <- Madoka[1:5,]; print(Smallmad)
```

I then wanted to change one of the columns in Smallmad, but something perculiar happaned.  
Any input to [5,3] other than the original input, “00:10:39” would result in an invalid factor level, and generate an \<NA\>. If I changed it back to the original input, it worked
properly again.  
Not-pictured below: I thought that maybe the data format being ‘factor’ meant that it “factored” somehow with the other time values. I don’t expect that the program is smart enough to recognize that “xx:xx:xx” is a time format without being specifically told that it *is* a time format, but I still tried changing it so it didn’t technically contradict any timestamps I could see. However, even changing [5,3] to “00:10:40” resulted in \<NA\>.

``` {r small2}
Smallmad[5,3] <- "The angriest Madoka"; print(Smallmad); Smallmad[5,3] <- "00:10:39";
```
I tried some additional commands with Smallmad, such as difftime which would result the mean difference in time for all the X values (i.e. how long the actual song is) but the result didn't make much sense.  From this I infer the data would need to be converted into a real time format.
```{dif}
mean.difftime(Smallmad[,6]); #Time difference of 11.8 
```
Turning Smallmad into a histogram results that it's not numeric data, either.  At a different time, I will return to Smallmad and figure out how to convert the characters into a real formatted timestamp, but for now it's time to move on and go forward with learning more from the book.